Quality of Service with Software Defined Network.

### 2017-03-06 report/paper.tex

\section{Introduction}\label{sec:intro}
Software-defined networking (SDN) is way to manage networks that separates the control plane from the forwarding plane.
SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brains” of the network.
The SDN Controller relays information to switches and routers via southbound APIs, and to the applications with northbound APIs.
One of the most well-known protocols used by SDN Controllers is OpenFlow, however, it isn’t the only SDN standard, but it is widely accepted.
Centralized, programmable SDN environments can easily adjust to the rapidly changing needs of businesses. SDN can lower costs and limit
wasteful provisioning, as well as provide flexibility and innovation for networks.
\#TODO da completare
  \subsection{OpenFlow}
  During 2006 thanks to M. Casado was developed Ethane a new network architecture on this conncept during 2011 was bord the Open Network Foundatin.
  This is a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
  The first version of OpenFlow is 1.0.0 release on 2009 but the last standardized version is v1.5.1 approved on march 2015 \cite{ONF_report}.
  \#TODO da sistemare
  \subsection{The state of the art}
  \#TODO stato dell arte e paper di riferimento  matteo

@@

    \subsubsection{altre}
    \#TODO andrea ?!?!?

@@

\section{Conclusions}\label{sec:conclusion}
Conclusions are a superbrief summary of what has been done and highlighting of the "take home message"

\#TODO

\begin{thebibliography}{99}
	\bibitem{pox} POX \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}
	\bibitem{openflow} OpenFlow \url{http://flowgrammable.org/sdn/openflow/}
	\bibitem{pox_repo} POX's repository \url{https://github.com/noxrepo/pox}
	\bibitem{VM} Virtual Machine \url{http://sdnhub.org/tutorials/pox/}
	\bibitem{router_board} Router Board \url{https://www.senetic.it/product/RB450G}
	\bibitem{netoworkx} NetworkX python's library \url{https://networkx.github.io/}
	\bibitem{ONF} Open Networking Foundation \url{https://www.opennetworking.org/sdn-resources/technical-library}
	\bibitem{ONF_report} v1.5.1 OpenFlow report \url{https://www.opennetworking.org/sdn-resources/technical-library}
	\bibitem{routerboard_doc} RouterBOARD 450G MikroTik documentation \url{http://www.routerboard.sk/files/pdf/rb450g_manual.pdf}
	\bibitem{routeboard_software} RouterBOARD MikroTik Software \url{https://mikrotik.com/software}
\end{thebibliography}

>>>>

\section{Introduction}\label{sec:intro}
Software-defined networking (SDN) is way to manage networks that separates the control plane from the forwarding plane.
SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brains” of the network.
The SDN Controller relays information to switches and routers via southbound APIs, and to the applications with northbound APIs.
One of the most well-known protocols used by SDN Controllers is OpenFlow, however, it isn’t the only SDN standard, but it is widely accepted.
Centralized, programmable SDN environments can easily adjust to the rapidly changing needs of businesses. SDN can lower costs and limit
wasteful provisioning, as well as provide flexibility and innovation for networks.
\#TODO da completare
  \subsection{OpenFlow}
  During 2006 thanks to M. Casado was developed Ethane a new network architecture on this conncept during 2011 was bord the Open Network Foundatin.
  This is a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
  The first version of OpenFlow is 1.0.0 release on 2009 but the last standardized version is v1.5.1 approved on march 2015 \cite{ONF_report}.
  \#TODO da sistemare
  \\
  \\
  \textbf{Openflow 1.0 messages}
Openflow defines a number of messages between a controller and a switch. Now there will be presented some messages we exploited:\cite{openflow}
\begin{itemize}

\item \emph{Hello} \newline This message is sent both from the controller and the switch who is looking for a controller after it
	is turned on. If the switch sent the message, the controller, after receiving it, will send the FeatureReq packet to that switch.
\item \emph{FeatureReq} \newline this is the message that the controller sends to a switch to know its abilities. The switch
	respond with a FeatureRes and when this last message reaches the pox controller, the ConnectionUp event is raised.
	With the featureRes the controller will know how many ports a switch has and their properties (e.g. if they are connected, link speed,...), the flow table capabilities and the supported statistics.
\item \emph{PortStatus} \newline this message is sent from a switch to the controller, to notify that some ports has changed their status.
		It's useful to check if a port has someone connected
\item \emph{StatsReq} \newline The controller will send this message to request the statistics. The payload of the message contains the type of statistic the controller want to know.
		The switch then will answer with the Stats reply, if it supports that type of stat. It will send a Bad_Request message if not.
\item \emph{PacketIn} \newline This message is sent from the switch to the controller whenever a packet arrives and it doesn't match
		any rule. The controller can decide to ignore that type of message, insert this message in a PacketOut message, or set
		a flow rule for packets like this using a FlowMod.
\item \emph{PacketOut} \newline This message is sent from the controller to forward a packet, which is the payload of the PacketOut message, to some
		port(s).
\item \emph{FlowMod} \newline This message set, delete or modify a rule inside a switch. Is the message that allows routing.
\\
\\
\end{itemize}

\textbf{Forwarding table:}f
\begin{itemize}
 \item \texttt{match fields:} sono i valori con cui i pacchetti vengono confrontati per cercare corrispondenze. Contengono la porta d'ingresso, l'intestazione del pacchetto, 
 e altri campi opzionali;
 \item \texttt{priority:} precedenza nel caso di voci con pi\`u corrispondenze;
 \item \texttt{counter:} incrementato ogni qualvolta che viene riscontrata la corrispondenza con un pacchetto;
 \item \texttt{istructions:} istruzioni da aggiungere all'\emph{Action-Set} o modifiche al flusso della pipeline;
 \item \texttt{timeouts:} massimo tempo per cui lo switch tiene in memoria questa voce, altre il quale la voce viene scartata;
 \item \texttt{cookie:} valori utilizzati per statistiche riguardo la singola voce, quindi non vengono considerati nell'elaborazione dei pacchetti dati;
 \item \texttt{flags:} utilizzate per differenziare la gestione della voce;
 \\
 \\
\end{itemize}

\textbf{Match Constraints}
principal match method that we can used with OpenFlow v1.0:
\begin{itemize}
 \item \texttt{in_port}: the port of swith;
 \item \texttt{dl_src/dl_dst}: 802.3 MAC source and destination of packet;
 \item \texttt{dl_type}: 802.3 EtherType;
 \item \texttt{nw_tos}: IPv4 ToS;
 \item \texttt{nw_proto}: IPv4 Protocol;
 \item \texttt{nw_src/nw_dst}: IPv4 Source/Destination;
 \item \texttt{tp_src/tp_dst}: TCP/UDP Source/Destination Port;
 \\
 \\
\end{itemize}

\textbf{Action}
With the OpenFlow v1.0 in practice we can do simply things:
\begin{itemize}
 \item \texttt{Output}: set output port;
 \item \texttt{SetDLSrc/Dst}: modify packet MAC address;
 \item \texttt{SetNWSrc/Dst}: modify packet IPv4 address;
 \item \texttt{SetNWTos}
 \item \texttt{SetTPSrc/Dst}
\end{itemize}
With from OpenFlow v1.1 in poi we can modify Time To Live and other things.

  \subsection{Related Work}
  The work presented in 
"Efficient topology discovery in OpenFlow-based Software Defined Networks" by F. Pakzad \cite{farzaneh} 
has been implemented to improve the controller speed and efficiency as it make it more reactive.
The introduction to the statistics power in a SDN network using the OpenFlow protocol has been found 
in the paper called "Getting traffic statistics from network devices in an SDN environment using OpenFlow" by D.J. Hamad \& others. [ put note here]
For traffic analysis we haven't found any paper describing the characteristics of
some types of flows, so we used our PCs to sniff the traffic. Traffic analysis will be discussed in section.
  \#TODO stato dell arte e paper di riferimento  matteo

@@

@@

\section{Conclusions}\label{sec:conclusion}
Conclusions are a superbrief summary of what has been done and highlighting of the "take home message"
ehi
\#TODO

\begin{thebibliography}{99}
	\bibitem{pox} POX \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}
	\bibitem{openflow} OpenFlow \url{http://flowgrammable.org/sdn/openflow/}
	\bibitem{pox_repo} POX's repository \url{https://github.com/noxrepo/pox}
	\bibitem{VM} Virtual Machine \url{http://sdnhub.org/tutorials/pox/}
	\bibitem{router_board} Router Board \url{https://www.senetic.it/product/RB450G}
	\bibitem{netoworkx} NetworkX python's library \url{https://networkx.github.io/}
	\bibitem{ONF} Open Networking Foundation \url{https://www.opennetworking.org/sdn-resources/technical-library}
	\bibitem{ONF_report} v1.5.1 OpenFlow report \url{https://www.opennetworking.org/sdn-resources/technical-library}
	\bibitem{routerboard_doc} RouterBOARD 450G MikroTik documentation \url{http://www.routerboard.sk/files/pdf/rb450g_manual.pdf}
	\bibitem{routeboard_software} RouterBOARD MikroTik Software \url{https://mikrotik.com/software}
	\bibitem{farzaneh}  Efficient Topology Discovery in Software Defined Networks F. Pakzad, M. Portmann, W. L. Tan and J. Indulska 
	School of ITEE, The University of Queensland Brisbane, Australia \url{https://www.researchgate.net/publication/271508903_Efficient_Topology_Discovery_in_OpenFlow-based_Software_Defined_Networks}
\end{thebibliography}
